---
title: "Spatial Analysis of the Flat Market in Prague"
subtitle: "Prostorové modely na Pražském trhu nemovitostí"
author: Petr Hrobař
output:
  github_document:
    pandoc_args: --webtex
  highlight: pygments
bibliography: Citace.bib
always_allow_html: true
editor_options: 
  chunk_output_type: console
---
  
```{r set knitr options, echo = F}
# Defaultní nastavení
# options(warn = -1)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width=12, fig.height=10)
```

# Abstract

Our study aims to examine the effect of spatial dependency in the Prague flat
market. Following the Tobler’s first law of geography: near things are more related
than distant things, we allow the flat’s price to be not only the function of its own
characteristics but also function of its neighbourhood unit characteristics. We also
evaluate different spatial dependence matrices for analysis of estimated results. The
findings are as follows. First, based on the positive parameter of spatial autocorrelation,
we confirm spatial dependency in our dataset, i.e. prices of flats tend to form spatial
clusters. Second, once controlling for spatial dependency, we are able to evaluate
flat’s prices more accurately. Third, based on the residual distribution across space,
we identify “grandiose“ clusters in which the price of estate can be multiple times
higher than outside the cluster simply due to the location factor.

# Models used

## Linear Regression
First model used in our study is simple linear regression given as:

$$ 
    y = X\beta + \varepsilon,
$$

This model does not take spatial dependency into account. In order to partially allow for spatial dependency, we propose simple feature based on statistical learning clustering algorithms. Since we are working with spatial data, we have coordinates available although we do not recommend including coordinates into linear model directly. 


## Spatial Lag Model

Spatial lag model can be obtained from when $\theta$ and $\lambda$ are both equal to 0. Under this assumption derived model allows for accounting for spatial dependency in dependent variable of neighbourhood units. Therefore, model can be written as


$$  y = \rho W y + X\beta + \varepsilon $$


## Spatial Error model

Second spatial model used in our study is spatial error model. Using this model specification we can account for spatial interactions among the error terms. Formal model form can be described as:


$$y = X \beta + u $$
$$u = \lambda W u + \varepsilon.$$


Selection for model can imply that some (spatially distributed) independent variable was not included in the model. Once again, parameters $\beta$, $\lambda$ are estimated by the ML, see e.g. @lesage2009introduction. Unlike the spatial lag model, coefficients from spatial error model can be directly interpreted as marginal effects. 


# Dataset

Flats estates are retrived from Czech estates site Sreality.cz which contains various estates that are available to rent
or buy. We suppose that Prague flats listed here are credible representation of the real flats market of the city, i.e.
the listed estates follow the same DGP as the estates not listed and thus not included in the dataset

Example of dataset can be inspected down below:

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)

df <- readr::read_csv("../Data and Scrapping/Dataset_Filtered_cleaned.csv")

df %>% 
  dplyr::select(price, Meters, Rooms, Mezone, KK, panel, balcony_or_terrase, novostavba) %>%
  head() %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```


# Empirical results

All estimated models can be inspected down below:




# References